---
title: "Text Processing / Feature Engeneering"
author: "Petro Tolochko"
date: ""
output:
  html_document
---

### Required Packages

```{r, message=FALSE, results='hide'}
library(tm)
library(tidyverse)
library(openNLP)
library(ggthemes)
library(ggrepel)
```


# Stop words and other preprocessing effects

```{r}
getwd()
setwd("~/Desktop/teaching/CEU_text_as_data/meeting_4/") # this is an example, paste your path here
```


Load the data:

```{r}

federalist <- read_csv("../meeting_3/federalist.csv")

```

### Basic Pre-processing

```{r}

clean_federalist <- federalist %>%
  mutate(                           # the mutate() function is part of dplyr package / allows to change stuff within the dataframe easily
    text   = str_to_lower(text),                # turn all letters to lowercase
    text   = str_replace_all(text, "\n", " "),  # replace '/n' carriage return symbols
    text   = str_remove_all(text, "[:punct:]"), # remove all punctuation
    man    = str_count(text, "\\Wman "),        # Basic regex (more about it later in the course. '\\W' part means at the begging of the word) and count those up
    by     = str_count(text, "\\Wby "),         # same
    upon   = str_count(text, "\\Wupon ")        # same
  ) %>%
  rowwise() %>%                                 # make future functions work rowwise
  mutate(
    length = length(str_split(text, " ")[[1]])  # calculate the length of the text (in words)
  )

```


What are the most used words in the federalist papers?

```{r}
dtm_federalist <- DocumentTermMatrix(clean_federalist$text,
                                     control = list(removePunctuation = TRUE,
                                         stopwords = FALSE))
dtm_federalist <- dtm_federalist %>% as.matrix()

dim(dtm_federalist)

most_frequent <- dtm_federalist %>% colSums()


most_frequent_df <- most_frequent %>% as.list() %>% as_tibble() %>%
  pivot_longer(everything())

most_frequent_df %>% arrange(desc(value))

```

```{r}

most_frequent_df %>%
  mutate(label = ifelse(value > 1200, name, NA)) %>%
  ggplot() +
  geom_point(aes(reorder(name, -value),  value)) +
  geom_text_repel(aes(reorder(name, -value),  value, label = label),
                  max.overlaps = 20) +
  theme_tufte() +
  ylab("Frequency") +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks = element_blank()
  )

```

Zipf's Law (for many types of data studied in the physical and social sciences, the rank-frequency distribution is an inverse relation):

```{r}


most_frequent_df %>%
  arrange(desc(value)) %>%
  mutate(rank = 1:nrow(.)) %>%
  ggplot(aes(rank, value)) +
  geom_line() +
  ylab("Frequency") +
  xlab("Rank") +
  scale_x_log10() +
  scale_y_log10() +
  theme_tufte()

```



Not amazing.

What if we remove the "basic" stop-words?

```{r}
dtm_federalist <- DocumentTermMatrix(clean_federalist$text,
                                     control = list(removePunctuation = TRUE,
                                         stopwords = TRUE))
dtm_federalist <- dtm_federalist %>% as.matrix()

dim(dtm_federalist)

most_frequent <- dtm_federalist %>% colSums()

most_frequent_df <- most_frequent %>% as.list() %>% as_tibble() %>%
  pivot_longer(everything())

most_frequent_df %>% arrange(desc(value))
```
Already looking a little better!

```{r}
most_frequent_df %>%
  mutate(label = ifelse(value > 500, name, NA)) %>%
  ggplot() +
  geom_point(aes(reorder(name, -value),  value)) +
  geom_text_repel(aes(reorder(name, -value),  value, label = label),
                  max.overlaps = 20) +
  theme_tufte() +
  ylab("Frequency") +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks = element_blank()

  )

```


```{r}

most_frequent_df %>%
  arrange(desc(value)) %>%
  mutate(rank = 1:nrow(.)) %>%
  ggplot(aes(rank, value)) +
  geom_line() +
  ylab("Frequency") +
  xlab("Rank") +
  scale_x_log10() +
  scale_y_log10() +
  theme_tufte()
```

# TFIDF

$$
W_{ij} \times log\frac{N}{n_j}
$$

Super easy, actually. There's already a function implemented in the `tm` package. It's called `weightTfIdf`, and we should pass it to the control param of the `DocumentTermMatrix` function:

```{r}
dtm_federalist <- DocumentTermMatrix(clean_federalist$text,
                                     control = list(removePunctuation = TRUE,
                                         stopwords = TRUE))

dtm_federalist_tfidf <- weightTfIdf(dtm_federalist)


dtm_federalist_tfidf <- dtm_federalist_tfidf %>% as.matrix()

dim(dtm_federalist_tfidf)

most_frequent <- dtm_federalist_tfidf %>% colSums()

most_frequent_df <- most_frequent %>% as.list() %>% as_tibble() %>%
  pivot_longer(everything())

most_frequent_df %>% arrange(desc(value))
```

Not bad... Let's plot this:

```{r}
most_frequent_df %>%
  mutate(label = ifelse(value > .107, name, NA)) %>%
  ggplot() +
  geom_point(aes(reorder(name, -value),  value)) +
  geom_text_repel(aes(reorder(name, -value),  value, label = label),
                  max.overlaps = 20) +
  theme_tufte() +
  ylab("tf-idf") +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks = element_blank()

  )

```


# Log Odds

First, let's see if `tfidf` can be a good discriminant function between `Hamilton` and `Madison`?

```{r}

dtm_federalist <- DocumentTermMatrix(clean_federalist$text,
                                     control = list(removePunctuation = TRUE,
                                         stopwords = TRUE))

dtm_federalist_tfidf <- weightTfIdf(dtm_federalist)

hamilton_ids <- which(clean_federalist$author == "HAMILTON")
madison_ids <- which(clean_federalist$author == "MADISON")

most_frequent_hamilton <- dtm_federalist_tfidf[hamilton_ids, ] %>%
  as.matrix() %>% colSums()
most_frequent_madison <- dtm_federalist_tfidf[madison_ids, ] %>%
  as.matrix() %>% colSums()



most_frequent_df_hamilton <- most_frequent_hamilton %>%
  as.list() %>%
  as_tibble() %>%
  pivot_longer(everything()) %>% 
  mutate(author = "HAMILTON")


most_frequent_df_madison <- most_frequent_madison %>%
  as.list() %>%
  as_tibble() %>%
  pivot_longer(everything()) %>% 
  mutate(author = "MADISON")

most_frequent_df <- bind_rows(most_frequent_df_hamilton,
                              most_frequent_df_madison)

hamilton_plot <- most_frequent_df %>%
  filter(author == "HAMILTON") %>%
  top_n(20, value) %>%
  ggplot() +
  geom_bar(aes(reorder(name, value), value),
           stat = "identity",
           fill = "steelblue") +
  coord_flip() +
  ylab("ti-idf") +
  theme_tufte() +
  theme(axis.title.y = element_blank())

madison_plot <- most_frequent_df %>%
  filter(author == "MADISON") %>%
  top_n(20, value) %>%
  ggplot() +
  geom_bar(aes(reorder(name, value), value),
           stat = "identity",
           fill = "indianred4") +
  coord_flip() +
  ylab("ti-idf") +
  theme_tufte() +
  theme(axis.title.y = element_blank())


cowplot::plot_grid(hamilton_plot, madison_plot,
                   ncol = 2)


```

Let's try to calculate log odds.

$$
logO^i_w = log\frac{f^i_w}{1 - f^i_w}
$$

```{r}
dtm_federalist <- DocumentTermMatrix(clean_federalist$text,
                                     control = list(removePunctuation = TRUE,
                                         stopwords = TRUE))

hamilton_ids <- which(clean_federalist$author == "HAMILTON")
madison_ids <- which(clean_federalist$author == "MADISON")

dtm_hamilton <- dtm_federalist[hamilton_ids, ] %>%
  as.matrix() %>% colSums()
freq_hamilton <- (dtm_hamilton + 1) / sum(dtm_hamilton + 1)  # laplace smoothing

dtm_madison <- dtm_federalist[madison_ids, ] %>%
  as.matrix() %>% colSums()

freq_madison <- (dtm_madison + 1) / sum(dtm_madison + 1)  # laplace smoothing


log_odds <- function(x) {
  log(x / (1 - x))
}

log_odds_ham <- log_odds(freq_hamilton)
log_odds_mad <- log_odds(freq_madison)


```

Now, log odds ratio:

$$
log\frac{O^i_w}{O^j_w} = log\frac{f^i_w}{1 - f^i_w}/\frac{f^j_w}{1 - f^j_w} = log\frac{f^i_w}{1 - f^i_w} - log\frac{f^j_w}{1 - f^j_w}
$$
```{r}
log_odds_ratio <- log_odds_ham - log_odds_mad



log_odds_ratio <- log_odds_ratio %>%
  as.list() %>%
  as_tibble() %>%
  pivot_longer(everything())

### Add word frequency ###

general_frequency <- dtm_federalist %>% as.matrix() %>% colSums() %>%
  as.list() %>%
  as_tibble() %>%
  pivot_longer(everything()) %>%
  mutate(freq = value) %>%
  select(-value)

log_odds_ratio <- log_odds_ratio %>%
  left_join(general_frequency)

log_odds_ratio %>%
  mutate(label = ifelse(value > 2 | value < -2, name, NA)) %>%
  ggplot() +
  geom_point(aes(freq, value, color = value)) +
  geom_text_repel(aes(freq, value, label = label),
                  max.overlaps = 50) +
  ylab("log odds ratio") +
  theme_tufte() +
  theme(legend.position = "none")


hamilton_plot <- log_odds_ratio %>%
  top_n(20, value) %>%
  ggplot() +
  geom_bar(aes(reorder(name, value), value),
           stat = "identity",
           fill = "steelblue") +
  coord_flip() +
  ylab("log odds ratio") +
  theme_tufte() +
  theme(axis.title.y = element_blank())

madison_plot <- log_odds_ratio %>%
  top_n(20, -value) %>%
  ggplot() +
  geom_bar(aes(reorder(name, -value), value),
           stat = "identity",
           fill = "indianred4") +
  coord_flip() +
  ylab("log odds ratio") +
  theme_tufte() +
  theme(axis.title.y = element_blank())


cowplot::plot_grid(hamilton_plot, madison_plot,
                   ncol = 2)

```


There are many problems with this method. Please refer to **Fightin’ Words: Lexical Feature Selection and Evaluation for Identifying the Content of Political Conflict** by **Monroe**, **Colaresi** & **Quinn** (2009) for a very cool overview and improvement on the method.
